Practical Machine Learning Project 
=========

# Project Summary

The goal of your project is to predict the manner in which they did the exercise which is that accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

```{r}
require(dplyr)
require(caret)
require(randomForest)

set.seed(1234)
```

## Processing data cleaned

Data in training and testing files loaded.

- mpl_training is the training data set
- mpl_testing is the test set

```{r}
mpl_training <- read.csv("/Users/truongpham/Documents/Coursera/8 Practical Machine Learning/data/pml-training.csv")
mpl_testing <- read.csv("/Users/truongpham/Documents/Coursera/8 Practical Machine Learning/data/pml-testing.csv")
```

Some varaiables is near to zero which make them useless in prediction, so they are removed from our data.

```{r}
zerodata <- nearZeroVar(mpl_training[,-160])
mpl_training_cleaned <- mpl_training[, -zerodata]
mpl_testing_cleaned <- mpl_testing[, -zerodata]
```

Some missing data shouldn't be included in predicted data. Therefore, they should be removed too.

```{r}
nas <- colSums(is.na(mpl_training_cleaned)) > 0
mpl_training_cleaned <- mpl_training_cleaned[, !nas]
mpl_testing_cleaned <- mpl_testing_cleaned[, !nas]
```

The first six variables in data looks unuseful in prediction, so they are deleted in dataset. They are X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, and num_window.

```{r}
mpl_training_cleaned <- mpl_training_cleaned[, 7:59]
mpl_testing_cleaned <- mpl_testing_cleaned[, 7:59]
```

Now, prediction is conducted.

## Cross-validation

Training dataset is divided into two parts of 70% and 30% for validationg.

```{r}

in_train <- createDataPartition(mpl_training_cleaned$classe, p = 0.7, list=FALSE)
training <- mpl_training_cleaned[in_train, ]
testing <- mpl_training_cleaned[-in_train, ]

```

## Prediction

Cross-validation uses four partions with method of random forest.


```{r, cache=TRUE}
fit <- train(classe ~ ., data=training, method="rf", trControl = trainControl(method = "cv", number = 4))
test_pred <- predict(fit, newdata=testing)
```

Cross-tabulation will be used for comparision of observed and predicted classes.


```{r}
confusionMatrix(test_pred, testing$classe)
```

The model accuracy is 99.46%, hence the error of ~ 0.54%. 

Prediction of test is followed as below.

```{r}
actual_predict <- predict(fit, newdata= mpl_testing_cleaned)
```

Ouput will be streamlized into file.

```{r}

# function for writing the output files
write_file = function(x){
  n = length(x)
  for(i in 1:n){
    fname = paste0("/Users/truongpham/Documents/Coursera/8 Practical Machine Learning/data/problem_id_",i,".txt")
    write.table(x[i],file=fname,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


write_file(actual_predict)
```

There are 20 files created for prediction of 20 sample tests.
